### 2014-05-20 sprout_eval.tex
Update codelcomp
Finish Sprout section?
Make Leslie fixes through end of Sprout

We are interested in performance metrics appropriate for a real-time
interactive application. In our evaluation, we report the
\emph{average throughput} achieved and the 95th-percentile
\emph{self-inflicted delay} incurred by each protocol, based on
measurement at the Cellsim.

@@

The {\em self-inflicted delay} is a lower bound on the end-to-end
delay that must be experienced between a sender and receiver, given
observed network behavior. We define it as follows: At any point in
time, we find the most recently-sent packet to have arrived at the
receiver. The amount of time since this packet was sent is a lower
bound on the instantaneous delay that must exist between the sender's
input and receiver's output in order to avoid a gap in playback or
other glitch at this moment. We calculate this instantaneous delay for
each moment in time. The 95th percentile of this function (taken over the
entire trace) is the amount of delay that must exist between the input
and output so that the receiver can recover 95\%~of the input signal
by the time of playback. We refer to this as ``95\%~end-to-end
delay.''

@@

\footnotesize
\def\svgwidth{\columnwidth}\input{varySprout2.pdf_tex}

@@

%\begin{center}
{
\small
\noindent \begin{tabular}{|l|l|l|l|}
\hline
& Direct & via Sprout & Change \\
\hline
\hline
Cubic throughput & 8336 kbps & 3776 kbps & \cellcolor{red!20}$-55\%$ \\
Skype throughput & 78 kbps & 490 kbps & \cellcolor{blue!20}$+528\%$ \\
Skype 95\% delay & 6.0~s & 0.17~s & \cellcolor{blue!20}$-97\%$ \\
\hline
\end{tabular}
}
\vspace{\baselineskip}

The results suggest that interactive applications can be
greatly aided by having their traffic run through Sprout along
with bulk transfers. Without Sprout to mediate, Cubic squeezes
out Skype and builds up huge delays. However, Sprout's conservatism about
delay also imposes a substantial penalty to Cubic's throughput.

>>>>

We are interested in performance metrics appropriate for a real-time
interactive application. In our evaluation, we report the
\emph{average throughput} achieved and the 95th-percentile
\emph{self-inflicted delay} incurred by each protocol, based on
measurement of packet entry and exit from the Cellsim.

@@

The {\em self-inflicted delay} is a lower bound on the end-to-end
delay that must be experienced between a sender and receiver, given
observed network behavior. We define it as follows: At any point in
time, we find the most recently sent packet to have arrived at the
receiver. The amount of time since this packet was sent is a lower
bound on the instantaneous delay that must exist between the sender's
input and receiver's output in order to avoid a gap in playback or
other glitch at this moment. We calculate this instantaneous delay for
each moment in time. The 95th percentile of this function (taken over the
entire trace) is the amount of delay that must exist between the input
and output so that the receiver can recover 95\%~of the input signal
by the time of playback. We refer to this as ``95\%~end-to-end
delay.''

@@

%\footnotesize
\def\svgwidth{\columnwidth}\input{varySprout2.pdf_tex}

@@

\begin{center}
{
\small
\noindent \begin{tabular}{|l|l|l|l|}
\hline
& Direct & via Sprout & Change \\
\hline
\hline
Cubic throughput & 8336 kbps & 3776 kbps & \cellcolor{red!20}$-55\%$ \\
Skype throughput & 78 kbps & 490 kbps & \cellcolor{blue!20}$+528\%$ \\
Skype 95\% delay & 6.0~s & 0.17~s & \cellcolor{blue!20}$-97\%$ \\
\hline
\end{tabular}
}
\end{center}

The results suggest that interactive applications can be greatly aided
by running their traffic---and any concurrent bulk transfers---through
Sprout. Without Sprout to mediate, Cubic squeezes out Skype and builds
up huge delays. However, Sprout's conservatism about delay
substantially reduces Cubic's throughput.

### 2014-05-20 sprout_impl.tex
Revise through some of eval
Make Leslie fixes through end of Sprout

For the device under test, the Verizon LTE and 1xEV-DO (3G) traces
used a Samsung Galaxy Nexus smartphone running Android 4.0. The AT\&T
trace used a Samsung Galaxy S3 smartphone running Android 4.0. The
T-Mobile trace used a Samsung Nexus S smartphone running Android 4.1.

@@

We then replay the traces in a network emulator, which we call
Cellsim (Figure~\ref{f:cellsim}). It runs on a PC and takes in packets
on two Ethernet interfaces, delays them for a configurable amount of
time (the propagation delay), and adds them to the tail of a
queue. Cellsim releases packets from the head of the queue to the
other network interface according to the same trace that was
previously recorded by Saturator. If a scheduled packet delivery
occurs while the queue is empty, nothing happens and the opportunity
to delivery a packet is wasted.\footnote{This accounting is done on a
  per-byte basis. If the queue contains 15 100-byte packets, they will
  all be released when the trace records delivery of a single
  1500-byte (MTU-sized) packet.}

@@

@@

\scriptsize
\noindent \begin{tabular}{|l|l|l|l|}
\hline
Program & Version & OS & Endpoints \\
\hline
\hline
Skype & 5.10.0.116 & Windows 7 & Core i7 PC \\
Hangout & as of 9/2012 & Windows 7 & Core i7 PC \\
Facetime & 2.0 (1070) & OS X 10.8.1 & MB Pro (2.3 GHz i7),\\
& & & MB Air (1.8 GHz i5) \\
TCP Cubic & in Linux 3.2.0 & & Core i7 PC \\
TCP Vegas & in Linux 3.2.0 & & Core i7 PC \\
LEDBAT & in $\mu$TP & Linux 3.2.0 & Core i7 PC \\
Compound TCP & in Windows 7 & & Core i7 PC \\
\hline
\end{tabular}
\end{figure}

>>>>

The Verizon LTE and 1xEV-DO (3G) traces
were collected with a Samsung Galaxy Nexus smartphone running Android 4.0. The AT\&T
trace used a Samsung Galaxy S3 smartphone running Android 4.0, and the
T-Mobile trace used a Samsung Nexus S smartphone running Android 4.1.

@@

We then replayed the traces in a network emulator, which we call
Cellsim (Figure~\ref{f:cellsim}). It runs on a PC and takes in packets
on two Ethernet interfaces, delays them for a configurable amount of
time (the propagation delay), and adds them to the tail of a
queue. Cellsim releases packets from the head of the queue to the
other network interface according to the same trace that was
previously recorded by Saturator. If a scheduled packet delivery
occurs while the queue is empty, nothing happens and the opportunity
to deliver a packet is wasted.\footnote{This accounting is done on a
  per-byte basis. If the queue contains 15 100-byte packets, they will
  all be released when the trace records delivery of a single
  1500-byte (MTU-sized) packet.}

@@

\begin{centering}

@@

%\scriptsize
\noindent \begin{tabular}{|l|l|l|l|}
\hline
Program & Version & OS & Endpoints \\
\hline
\hline
Skype & 5.10.0.116 & Windows 7 & Core i7 PC \\
Hangout & as of 9/2012 & Windows 7 & Core i7 PC \\
Facetime & 2.0 (1070) & OS X 10.8.1 & MB Pro (2.3 GHz i7),\\
& & & MB Air (1.8 GHz i5) \\
TCP Cubic & in Linux 3.2.0 & & Core i7 PC \\
TCP Vegas & in Linux 3.2.0 & & Core i7 PC \\
LEDBAT & in $\mu$TP & Linux 3.2.0 & Core i7 PC \\
Compound TCP & in Windows 7 & & Core i7 PC \\
\hline
\end{tabular}

\end{centering}
\end{figure}

### 2014-05-20 sprout_problem.tex
Make Leslie fixes through end of Sprout

Most (in fact, all, to our knowledge) deployed cellular wireless
networks enqueue each user's traffic in a separate queue. The base
station schedules data transmissions taking both per-user
(proportional) fairness and channel quality into
consideration~\cite{propfair}. Typically, each user's device is
scheduled for a fixed time slice over which a variable number of
payload bits may be sent, depending on the channel conditions, and
users are scheduled in roughly round-robin fashion. The isolation
between users' queues means that the dominant factor in the end-to-end
delay experienced by a user's packets is {\em self-interaction},
rather than cross traffic. If a user were to combine a high-throughput
transfer and a delay-sensitive transfer, the commingling of these
packets in the same queue would cause them to experience the same
delay distributions. The impact of other users on delay is
muted. However, competing demand can affect the throughput that a user
receives.

@@

Many cellular networks employ a non-trivial amount of packet
buffering. For TCP congestion control with a small degree of
statistical multiplexing, a good rule-of-thumb is that the buffering
should not exceed the bandwidth-delay product of the connection. For
cellular networks where the ``bandwidth'' may vary by two orders of
magnitude within seconds, this guideline is not particularly useful. A
``bufferbloated''~\cite{bufferbloat} base station at one link rate
may, within a short amount of time, be under-provisioned when the link
rate suddenly increases, leading to extremely high IP-layer packet
loss rates (this problem is observed in one
provider~\cite{Mahajan12}).

@@

Sprout is not a traditional congestion-control scheme, in that its
focus is directed at adapting to varying link conditions, not to
varying cross traffic that contends for the same bottleneck queue. Its
improvements over existing schemes are found when queueing delays are
imposed by one user's traffic. This is typically the case
when the application is running on a mobile device, because cellular
network operators generally maintain a separate queue for each
customer, and the wireless link is typically the bottleneck.  An
important limitation of this approach is that in cases where these
conditions don't hold, Sprout's traffic will experience the same
delays as other flows.
%We believe this is an unavoidable consequence
%of an end-to-end approach.

>>>>

Most (in fact, all, to our knowledge) deployed cellular wireless
networks enqueue each user's traffic in a separate queue. The base
station schedules data transmissions taking both per-user
(proportional) fairness and channel quality into
consideration~\cite{propfair}. Typically, each user's device is
scheduled for a time slice or resource block in which a variable number of
payload bits may be sent, depending on the channel conditions, and
users are scheduled in roughly round-robin fashion. The isolation
between users' queues means that the dominant factor in the end-to-end
delay experienced by a user's packets is {\em self-interaction},
rather than cross traffic. If a user were to combine a high-throughput
transfer and a delay-sensitive transfer, the commingling of these
packets in the same queue would cause them to experience the same
delay distributions. The impact of other users on delay is
muted. However, competing demand can affect the throughput that a user
receives.

@@

Many cellular networks employ a non-trivial amount of packet
buffering. For TCP congestion control with a small degree of
statistical multiplexing, a good rule of thumb is that the buffering
should not exceed the bandwidth-delay product of the connection. For
cellular networks where the ``bandwidth'' may vary by two orders of
magnitude within seconds, this guideline is not particularly useful. A
``bufferbloated''~\cite{bufferbloat} base station at one link rate
may, within a short amount of time, be under-provisioned when the link
rate suddenly increases, leading to extremely high IP-layer packet
loss rates (this problem is observed in one
provider~\cite{Mahajan12}).

@@

Sprout is not a traditional congestion-control scheme, in that its
focus is directed at adapting to varying link conditions, not to
varying cross traffic that contends for the same bottleneck queue. Its
improvements over existing schemes are found when queueing delays are
imposed by the user's own traffic.% This is typically the case
%when the application is running on a mobile device, because cellular
%network operators generally maintain a separate queue for each
%customer, and the wireless link is typically the bottleneck.  An
%important limitation of this approach is that in cases where these
%conditions don't hold, Sprout's traffic will experience the same
%delays as other flows.
%We believe this is an unavoidable consequence
%of an end-to-end approach.

### 2014-05-20 sprout_sprout.tex
Make Leslie fixes through end of Sprout

In our measurements, we found that estimating the capacity (by which
we mean the maximum possible bit rate or throughput) of cellular links
is challenging, because they do not have a directly observable rate
{\em per se}. Even in the middle of the night, when average throughput
is high and an LTE device may be completely alone in its cell, packet
arrivals on a saturated link do not follow an observable
isochronicity. This is a roadblock for packet-pair techniques
(\cite{packetpair}) and other schemes to measure the available
throughput.

@@

Sprout needs to be able to estimate the link speed, both now and in
the future, in order to predict how many packets it is safe to send
without risking their waiting in a network queue for too long. An
uncertain estimate of future link speed is worth more caution than a
certain estimate, so we need to quantify our uncertainty as well as
our best guess.

We therefore treat the problem in two parts. We model the link and
estimate its behavior at any given time, preserving our full
uncertainty. We then use the model to make forecasts about how many
bytes the link will be willing to transmit from its queue in the near
future. Most steps in this process can be precalculated at program
startup, meaning that CPU usage (even at high throughputs) is less
than 5\% of a current Intel or AMD PC microprocessor. We have not
tested Sprout on a CPU-constrained device or tried to optimize it
fully.

@@

These steps constitute Bayesian updating of the probability
distribution on the current value of $\lambda$.

One important practical difficulty concerns how to deal with the
situation where the queue is underflowing because the sender has not
sent enough. To the receiver, this case is indistinguishable from an
outage of the service process, because in either case the receiver
doesn't get any packets.

@@

\subsection{The control protocol}

>>>>

In our measurements, we found that estimating the capacity (by which
we mean the maximum possible bit rate or throughput) of cellular links
is challenging, because they do not have a directly observable rate
{\em per se}. Even in the middle of the night, when average throughput
is high and an LTE device may be completely alone in its cell, packet
inter-arrival intervals on a saturated link are highly variable. This
is a roadblock for packet-pair techniques (\cite{packetpair}) and
other schemes to measure the available throughput.

@@

Sprout needs to be able to estimate the link speed, both now and in
the future, in order to predict how many packets it is safe to send
without risking their waiting in a network queue for too long. An
uncertain estimate of future link speed deserves more caution than a
precise estimate, so we quantify our uncertainty as well as
our best guess.

We therefore treat the problem in two parts. We model the link and
estimate its behavior at any given time, preserving the full
uncertainty. We then use the model to make forecasts about how many
bytes the link will be willing to transmit from its queue in the near
future. Most steps in this process can be precalculated at program
startup, meaning that CPU usage (even at high throughputs) is less
than 5\% of a current Intel or AMD PC microprocessor. We have not
tested Sprout on a CPU-constrained device or tried to optimize it
fully.

@@

Steps 2 and 3 constitute Bayesian updating of the probability
distribution on the current value of $\lambda$.

One challenge concerns the
situation where the queue is underflowing because the sender has not
sent enough. To the receiver, this case is indistinguishable from an
outage of the service process, because in either case the receiver
doesn't get any packets.

@@

\subsection{Feedback from the receiver to sender}
